---
title: "Data Analysis Exercise"
subtitle: "BEAM Dashboard"
author: Seth Lattner
editor: visual
output:
  html_document:
    toc: FALSE
---

### About the Data

This [dataset](https://data.cdc.gov/Foodborne-Waterborne-and-Related-Diseases/BEAM-Dashboard-Report-Data/jbhn-e8xn) is part of the **BEAM (Bacteria, Enterics, Amoeba, and Mycotics) Dashboard**, collected by the System for Enteric Disease Response, Investigation, and Coordination (SEDRIC). It contains information on food- and water-borne pathogens circulating in the United States.

### Loading Data

```{r, message=FALSE, warning=FALSE}
#load required packages
library(tidyverse)
library(here)
library(zoo)
```

```{r}
#load data
BEAM_data <- read_csv(here("dataanalysis_exercise", "data", "BEAM_Dashboard_-_Report_Data.csv"))
glimpse(BEAM_data)
```

### Data Cleaning

Now that we have the data loaded, let's tidy it up to get it ready for analysis. Thankfully, this is a relatively clean dataset, so wrangling it shouldn't be too difficult.

First, I'm going to reclassify the categorical variables.

```{r}
BEAM_data <- BEAM_data %>%
  mutate(State = as.factor(State)) %>%
  mutate(Source = as.factor(Source)) %>%
  mutate(Pathogen = as.factor(Pathogen)) %>%
  mutate(`Serotype/Species` = as.factor(`Serotype/Species`))

glimpse(BEAM_data)  
```

I also want to group the total number of isolates of each pathogen for each year. This will help simplify the results and can be used for further analysis.

```{r}
grouped_data <- BEAM_data %>%
  group_by(Year, Pathogen) %>%
  summarise(sum_isolates = sum(Number_of_isolates), .groups = 'drop')
glimpse(grouped_data)
```

Now that these variables are reclassified, I want to change the NA's in a few columns to 0's, which is more indicative of their actual value.

```{r}
BEAM_data_clean <- BEAM_data %>%
  replace(is.na(BEAM_data), 0)

glimpse(BEAM_data_clean)
```

That will conclude the data cleaning I will be conducting. Perhaps more wrangling could have been done, but I think this is a good start and sets up some the next contributor fairly well.

```{r}
#save processed data
BEAM_data_clean %>% saveRDS(here::here("dataanalysis_exercise", "data", "BEAM_data_clean.rds"))
```

### Exploring the Data

Now that the data is clean, I want to do some quick exploration of it before handing it off to the next contributor for a more complete analysis.

The following plots will help give a simple overview of what the data looks like without getting into any modeling.

```{r, message= FALSE, warning=FALSE}
ggplot(BEAM_data_clean, aes(Pathogen, Number_of_isolates, color=Pathogen))+
  geom_col()+
  labs(title = "Total Number of Isolates from Each Pathogen Type")+
  xlab("Pathogen Type")+
  ylab("Number of Isolates")+
  theme_classic()+
  theme(legend.position = "none")

ggplot(grouped_data, aes(Year, sum_isolates, color=Pathogen))+
 geom_smooth()+
  labs(title = "Total Number of Isolates for Each Pathogen, 2017-2022")+
  xlab("Year")+
  ylab("Number of isolates")+
  theme_classic()
```

### Data Analysis

There are many interesting questions that can be answered with this dataset. Among these, I am interested in:

-   The relationship between source and pathogen type and/or number of isolates

-   The relationship between serovar/species and outbreak association and/or number of new multistate outbreaks

-   The impact of state/year on the number of new multistate outbreaks

However, this dataset could be used to address many other questions, and I will leave that to the next contributor's judgement. Looking forward to seeing how this data will be analyzed!
