---
title: "Flu Analysis"
subtitle: "Exploration"
author: "Seth Lattner"
output:
  html_document:
    toc: FALSE
---

This is the second file of a three part data analysis exercise, conducted on the dataset from McKay et al 2020, found [here](https://doi.org/10.5061/dryad.51c59zw4v). This file contains data exploration steps in preparation for further analysis.

#### Load Data/Packages

```{r, warning=FALSE, message=FALSE}
#load necessary packages
library(tidyverse)
library(ggplot2)
library(kableExtra)
library(here)
```

```{r}
#load and view data
flu_data <- readRDS(here::here("fluanalysis", "data", "flu_data_clean.RDS"))
glimpse(flu_data)
```

#### Explore Data

Let's take a look at some summary stats for this dataset.

```{r}
summary(flu_data)
```


The two main outcomes of interest are BodyTemp and Nausea. I will plot a number of variables against these outcomes to see if any trends are immediately noticeable. 

First, I want to look at histogram of body temperature to see the range and distribution.

```{r}
flu_data %>%
  ggplot()+
  geom_histogram(aes(BodyTemp))+
  theme_classic()
```

This shows that a majority of patients recorded temperatures in a normal (non-fever) range. A decent proportion had temperatures between 99-100 F, with fewer being above 100 F. Temperatures up to 103.1 F were recorded.

#### Body Temperature

Now I will look at a number of plots comparing body temperature vs some relevant categorical predictors.

Boxplot of body temperature vs reported fever symptoms.

```{r}
ggplot(flu_data, aes(BodyTemp, SubjectiveFever))+
  geom_boxplot(fill="gray90")+
  theme_classic()
```

It looks like in general, patients reporting fever symptoms had higher body temperatures than those that did not. The interquartile range for patients reporting fevers falls between normal temperature to low-grade fever. Several patients with body temperatures above 100 F did not report any fever symptoms.


Boxplot of body temperature vs reported chills/sweats.

```{r}
ggplot(flu_data, aes(BodyTemp, ChillsSweats))+
  geom_boxplot(fill="gray90")+
  theme_classic()
```

It seems that patients with higher body temperatures experienced chills/sweats slightly more commonly than those with lower body temperatures.

Boxplot of body temperature vs reported weakness. 

```{r}
ggplot(flu_data, aes(BodyTemp, Weakness))+
  geom_boxplot(fill="gray90")+
  theme_classic()
```

In general, higher body temperatures were recorded in patients experiencing severe weakness than those with moderate, mild, or no weakness. 

Boxplot of body temperature vs cough intensity (ranked).

```{r}
ggplot(flu_data, aes(BodyTemp, CoughIntensity))+
  geom_boxplot(fill="gray90")+
  theme_classic()
```

In general, higher body temperatures were recorded in patients experiencing severe cough intensity than those with moderate, mild, or no coughing.

Boxplot of body temperature vs myalgia. 

```{r}
ggplot(flu_data, aes(BodyTemp, Myalgia))+
  geom_boxplot(fill="gray90")+
  theme_classic()
```

In general, higher body temperatures were recorded in patients experiencing severe myalgia than those with moderate, mild, or no myalgia.

#### Nausea

Now I want to visualize nausea as predicted by a number of categorical variables.

Nausea vs reported abdominal pain
```{r}
ggplot(flu_data, aes(Nausea))+
  geom_histogram(aes(fill=AbPain), stat = "count")+
  theme_classic()
```

Abdominal pain was reported more often in patients experiencing nausea.

Nausea vs reported vomitting
```{r}
ggplot(flu_data, aes(Nausea))+
  geom_histogram(aes(fill=Vomit), stat = "count")+
  theme_classic()
```

Vomitting was reported more often in patients experiencing nausea.

Nausea vs reported diarrhea
```{r}
ggplot(flu_data, aes(Nausea))+
  geom_histogram(aes(fill=Diarrhea), stat = "count")+
  theme_classic()
```

Diarrhea was reported more often in patients experiencing nausea.

Nausea vs reported chills/sweats
```{r}
ggplot(flu_data, aes(Nausea))+
  geom_histogram(aes(fill=ChillsSweats), stat = "count")+
  theme_classic()
```

Chills/sweats were reported more often in patients experiencing nausea.