---
title: "Flu Analysis"
subtitle: "Fitting"
author: "Seth Lattner"
output:
  html_document:
    toc: FALSE
---

This is the final file of a three part data analysis exercise, conducted on the dataset from McKay et al 2020, found [here](https://doi.org/10.5061/dryad.51c59zw4v). This file contains data analysis and model fitting. 

#### Load Data/Packages

```{r, warning=FALSE, message=FALSE}
#load necessary packages
library(tidyverse)
library(tidymodels)
library(kableExtra)
library(here)
library(performance)
```

```{r, message=FALSE}
#load and view data
flu_data <- readRDS(here::here("fluanalysis", "data", "flu_data_clean.RDS"))
glimpse(flu_data)
```

### Model Fitting

#### Body Temperature

First, I want to fit a linear regression model using our first outcome of interest, body temperature, and a single predictor variable, runny nose.

```{r}
flu_lm1 <- linear_reg() %>%
  set_engine("lm") %>%
  fit(BodyTemp ~ RunnyNose, data = flu_data)
kableExtra::kable(tidy(flu_lm1)) %>%
  kableExtra::kable_classic(full_width = FALSE, font = "Arial")
glance(flu_lm1)
```

Based on the R-squared value and the RMSE, this does not seem to be a great fit to the data. I want to throw in all of the other variables into a global model and see if this does a better job of explaining the data than the first model.

```{r}
flu_lm2 <- linear_reg() %>%
  set_engine("lm") %>%
  fit(BodyTemp ~ . , data = flu_data)
kableExtra::kable(tidy(flu_lm2)) %>%
  kableExtra::kable_classic(full_width = FALSE, font = "Arial")
glance(flu_lm2)
```

Let's compare these models side-by-side.

```{r}
compare_performance(flu_lm1, flu_lm2)
```

Even though it is not a great fit to the data (R-squared = 0.13), the global model is better at explaining the data than the first model based on R-sq and RMSE (as well as AIC).

#### Nausea

Now, I want to fit a logistic regression model using our second outcome of interest, nausea, with a single predictor variable, runny nose.

```{r}
flu_glm1 <- logistic_reg() %>%
  set_engine("glm") %>%
  fit(Nausea ~ RunnyNose, data = flu_data)
kableExtra::kable(tidy(flu_glm1)) %>%
  kableExtra::kable_classic(full_width = FALSE, font = "Arial")
glance(flu_glm1)
```

Same as before, I want to fit a global model using nausea and all of the other variables in the dataset and compare it to the first model.

```{r}
flu_glm2 <- logistic_reg() %>%
  set_engine("glm") %>%
  fit(Nausea ~ . , data = flu_data)
kableExtra::kable(tidy(flu_glm2)) %>%
  kableExtra::kable_classic(full_width = FALSE, font = "Arial")
glance(flu_glm2)
```

Let's see which model is a better fit.

```{r}
compare_performance(flu_glm1, flu_glm2)
```

Once again, the global model was better at explaining the data than the univariate model. 

