---
title: "Flu Analysis"
subtitle: "Wrangling"
author: "Seth Lattner"
output:
  html_document:
    toc: FALSE
---

This is the first file of a three part data analysis exercise, conducted on the dataset from McKay et al 2020, found [here](https://doi.org/10.5061/dryad.51c59zw4v). This file contains data wrangling and cleaning steps in preparation for further analysis.

#### Load Packages/Data

```{r, warning=FALSE, message=FALSE}
#load necessary packages
library(tidyverse)
library(ggplot2)
library(here)
```


```{r}
#load in data
flu_raw <- readr::read_rds(here::here("fluanalysis", "data", "SympAct_Any_Pos.Rda"))

#view data
glimpse(flu_raw)
summary(flu_raw)
```

#### Cleaning/Wrangling

This dataset will require a small amount of cleaning - namely, I will remove some variables that will not be useful for future analysis and remove missing values.

```{r}
flu_data <- flu_raw %>%
  select(-contains(c("Score", "Total", "DxName", "Activity", "FluA", "FluB", "Unique.Visit"))) %>%
  drop_na()
glimpse(flu_data)
```

#### Save Cleaned Data

This completes the data cleaning that I will be doing, so I will now save the clean data as a new RDS file.

```{r}
flu_data %>%
  saveRDS(here::here("fluanalysis", "data", "flu_data_clean.RDS"))
```